---
title: "A proposed typology for project names"
author: |
    | Hari Subhash
    | Data Scientist @NRGI
output:
  html_notebook:
    code_folding: hide
    highlight: kate
    smart: yes
    theme: cosmo
  html_document:
    df_print: paged
date: '`r Sys.Date()`'
css: custom.css
---

```{r, message=FALSE, warning=FALSE, echo=FALSE}
rm(list = ls())
library(tidyverse); library(googlesheets); library(stringdist); library(ggthemes); library(scales)

cleanProjects <- read_csv("../../../resource-project-data/Data Pipeline/All Data Backups/Current website data/allProject.csv")

nonReconciledProjects <- gs_title("Project Cleaning") %>% 
    gs_read(ws = "Cannot be reconciled", verbose = F) %>% 
    filter(!is.na(rawName))

secondaryData <- gs_title("Final Secondary - Project") %>% 
    gs_read(ws = "Data", verbose = F)

projectVals <- nonReconciledProjects$rawName %>% str_replace_all(., " ", "") %>% str_to_lower()
d <- stringdistmatrix(projectVals, method = "osa")

hc <- hclust(as.dist(d))

projClusters <- nonReconciledProjects %>% 
    mutate(cluster = cutree(hc, h = 4)) %>% 
    group_by(cluster) %>% 
    mutate(nObs = n()) %>% 
    arrange(desc(nObs), cluster)

secondaryPaymentData <- left_join(secondaryData %>% select(1:3, 5, 6, 7), cleanProjects, by = c("projectCountry", "projectName"))
```

This piece summarises the different types of names used for projects by reporting companies in mandatory disclosures. I have proposed a loose typology that could be our starting point for how we want to categorize project names. The end goal is to narrow the number of projects that we cannot identify and expand the ones that we can collect data on.

###***A quick overview of the frequency of different types of project names in the data***

The charts below use different types that are not currently being used to categorize projects. Right now we have two categories, one that we can identify and collect data for and another for which we cannot identify or collect data. This is an attempt to expand the second category of projects that we do not collect data for so that we capture more "secondary" data on some of the projects that we currently don't cover.

The first chart shows the number of rows of data for each project category while the second shows the total payments for each. The main takeaway from these charts is that while unidentifed projects constutute about half the number of rows of payments data, they constitute only 10% of the total payments amounts. Another insight is that the corporate category constitutes a significant proportion of payments despite being a smaller number of rows in the data, i.e. the average payment for each row of Corporate category is very high.


```{r, fig.align='center', fig.width=12, fig.height = 4, echo=FALSE}
actualProject <- paste0(secondaryData[secondaryData$projectType != "Corporate", ]$projectCountry, secondaryData[secondaryData$projectType != "Corporate",] $projectName) %>% 
    str_to_lower() %>% 
    str_trim()
indentifiableCorp <- paste0(secondaryData[secondaryData$projectType == "Corporate", ]$projectCountry, secondaryData[secondaryData$projectType == "Corporate",] $projectName) %>% 
    str_to_lower() %>% 
    str_trim()


newProjects <- cleanProjects %>%
    mutate(combo = paste0(projectCountry, projectName)%>% str_to_lower() %>% str_trim()) %>% 
    mutate(nameType = case_when(
        combo %in% actualProject ~ "Identifiable Mine or Oil-Gas Asset/Collection",
        str_detect(str_to_lower(projectName), "alberta|saskatchewan|ab project|wyoming|dawson|daytona|new mexico") ~ "Generic Asset Collection",
        combo %in% indentifiableCorp ~ "Identifiable Corporate Entity",
        str_detect(str_to_lower(projectName), "corporate|head office|marketing|entity") ~ "Generic Corporate Entity",
        str_detect(str_to_lower(projectName), "explor|licens|shore|generati|closed|noncontractual") ~ "Generic Industry Term",
        str_detect(str_to_lower(projectName), "attribut|alloca|contract|other|no name|nonproject") ~ "Non Attributable",
        TRUE ~ "Unidentified"
    )) %>% 
    mutate(fillColor = case_when(
        nameType %in% c("Identifiable Corporate Entity", "Identifiable Mine or Oil-Gas Asset/Collection") ~ "Identifiable",
        TRUE ~ "Not identifiable"
    ))

p1 <- newProjects %>% 
    group_by(nameType) %>% 
    summarise(nObs = n(), fillColor = unique(fillColor)) %>% 
    ggplot(., aes(x = reorder(nameType, -nObs), y = nObs, fill = fillColor)) +
    geom_bar(stat = "identity", width = 0.8) +
    scale_x_discrete(labels = function(x) str_wrap(x, width = 16)) +
    scale_fill_manual(values = c("Identifiable" = "darkgreen", "Not identifiable" = "darkgrey")) +
    coord_flip() +
    labs(
        x = "Project Type", 
        y = 'No. of rows of payment data', 
        title = str_wrap("About half the payments rows are not associated with a recognizable name (either corporate or actual project)", width = 50)
        ) +
    theme_tufte() +
    theme(
        legend.position = "none",
        plot.title = element_text(size = 14, family = "Palatino"),
        axis.title = element_text(size = 12),
        axis.text = element_text(size = 11)
        )

p2 <- newProjects %>% 
    group_by(nameType) %>% 
    summarise(totPay = sum(projectPayment, na.rm = T)/1e9, fillColor = unique(fillColor))%>% 
    ggplot(., aes(x = reorder(nameType, -totPay), y = totPay, fill = fillColor)) +
    geom_bar(stat = "identity", width = 0.6) +
    scale_fill_manual(values = c("Identifiable" = "darkgreen", "Not identifiable" = "darkgrey")) +
    scale_x_discrete(labels = function(x) str_wrap(x, width = 16)) +
    scale_y_continuous(labels=dollar_format(prefix="$")) +
    coord_flip() +
    labs(x = "Project Type", y = 'Total Payment (in billion USD)', title = str_wrap("Only about 10% of the total amounts have unidentified names", width = 50),
         fill = "Current Project Category") +
    theme_tufte() +
    theme(
        plot.title = element_text(size = 14, family = "Palatino"),
        axis.title = element_text(size = 12),
        axis.text = element_text(size = 11)
        )

cowplot::plot_grid(p1, p2)
```



#### ***An explaination of the types***

I have defined these categories as follows (<span style="background-color:lightgreen">highlighted</span> ones are the types of projects that we currently collect data for)

1. <span style="background-color:lightgreen">Identifiable oil-gas or mining asset/collection</span>: By clearly identifiable we mean a project that can be identified solely based on the name of the country and the projects name. 
    + The ideal case is when a project refers to a single asset. For instance, `Agua Rica Project Minera Agua Rica Llc`, Argentina is a name used by Yamana Incorporated to report Exploration Fees made through their interest in the Agua Rica mine. While the name contains the corporate entity that manages the mine we were still able to tag it to a meaningful project for which we collected additional information on commodities (copper, molybdenum, gold, silver and magnetite) and location (-27.370454, -66.281453). 
    + The less ideal case is when the name refers to a collection of assets. For instance, `Thrace Basin`, Turkey reported by Valeura Energy Incorporated is a [clearly identified](https://en.wikipedia.org/wiki/Thrace_Basin) collection of oil and gas fields (400 wells). However, without the name of the specific block that they are referring to we won't be able to pinpoint the exact asset. As a result our secondary information for this collection is limited to an approximate location. In addition, exercises such as linking contracts will be harder to do.
2. <span style="background-color:lightgreen">Identifiable corporate entity</span>: In some cases, payments made by legal entities (often local subsidiaries paying taxes) cannot be attributed to specific projects (I am not sure why these payments cannot be attributed 👋**@MANTRA**). For instance, `Statoil Nigeria Limited` is a clearly identifiable local subsidiary. We categorize these as corporate entities and collect additional information (location, country of registration etc.)
3. Generic Industry Term: These include `exploration`, `license` etc. as references to projects. There are cases in which a company uses a reference to a specific project even when it is an exploration, for instance, Agua Rica (in case 1) is a exploration project, however, Yamana chose to use a specific name instead of a generic term such as `exploration`. This might be due to confidentiality reasons but I wonder if there is a better middle ground like using a collection name (👋**@MANTRA**).
4. Generic Asset Collection: These are cases were companies report the names of large geographic areas. For instance, `Alberta` Canada is an oft used term for projects which refers to an entire province in Canada that covers approx. 661,848 square kilometers in area. There is no meaningful way in which these can be tagged to a specific area or project. However, we could potentially infer that they are referring to an oil and gas field since that happens to be the primary natural resource in Alberta Canada. These seem to be clear cases of loosely applied reporting standards.
5. Generic Corporate Entity: Often instead of using the name of the specific legal entity that made a payment, companies use generic terms like `corporate`, `marketing offices`, `head office`, `entity level` etc. in reports (Do you know why this might be? 👋**@MANTRA**). This again seems to be an area in which we could advocate for improvement.
6. Non-attributable: Companies often use terms such as `non-attributable`, `non-allocable` etc to specify payments that could not be attributed to projects. These potentially fall under the "Generic Industry Term" or "Generic Corporate Term" categories. (👋**@MANTRA**)
7. Unidentified: These are cases were we have no clue what a company is referring to (without referring to their internal documents). These are cases were we couldn't find a public reference to an oil/gas or mining project that shares the name used by the company to report it.


👉**A few points to discuss**👈

1. Does this make sense?
2. There is some information in the categories that we are not currently collecting data on. For instance, a Generic industry term such as "Exploration" could be of value for a researcher if we are able to classify these accurately i.e. use a category instead of the raw name to describe the payment.
3. There are some types such as the Generic Asset Collection (the case were companies use a province, such as Alberta or a country name as the project) were it might worthwhile to think about advocating for better disclosure standards.


<!-- 👉**A short case for secondary data collection:**👈  -->
<!--  By "secondary", I mean any data that is beyond what is available in a mandatory disclosure report (payments made to govt. entities or projects). Secondary data collection currently includes data on projects, govt. entities and companies from mandatory disclosure reports that we are able to publicly identify. For instance, in the case of projects that we can identify we collect additional information such as type, commodity, location etc. -->

<!-- This secondary data collection has downstream implications for the development of the resource projects website for insance project/company/govt. entity pages require information that needs to be presented. It also enhances our capacity to put payment data to better use by providing contextual information that can inform our analysis. For instance, the ability to see dependence on certain commodities, tie payments to sub-national jurisdiction, study tax avoidance/base-erosion etc. We are also improving the linkages with third party data sources such SNL and Rystad, which would help justify those subscriptions and add richness to our analysis. Finally cleaned projects/govt entities and companies can allow us to link different datasets like EITI, RC etc to mandatory disclosure payments. Over the next 6 months I am hoping to spend some time developing a few analytical pieces that hopefully will showcase these use cases better 🤞🤞🤞. -->
