---
title: "Add data to pieline"
date: '`r Sys.Date()`'
output: 
  html_notebook: 
    number_sections: yes
    toc: true
    toc_depth: 2
    toc_float: true
---
##Introduction
This file explores different aspects of the payment data disclosed through ESTMA filings. It contains a series of charts and tables that describe different aspects of the data.

```{r, echo = FALSE, message=FALSE, results='hide'}
rm(list = ls())
library(tidyverse); library(knitr); library(kableExtra); library(ggthemes); library(stringr); library(viridis)
options(knitr.table.format = "html") 

##Load the data
sourceClean <- read_csv("../Payment Data Pipeline/Data/Pipeline/Clean Data/sourceClean.csv")
projectClean <- read_csv("../Payment Data Pipeline/Data/Pipeline/Clean Data/projectClean.csv")
entityClean <- read_csv("../Payment Data Pipeline/Data/Pipeline/Clean Data/entityClean.csv")
```

##Summary Data
This report compares the payments made to Mexico, Peru and Colombia. There are two sets of payment information to consider, project payments and entity payments. The table below shows the general characteristics of the data.

```{r, echo = FALSE, message=FALSE, results='asis'}
projectData <- projectClean %>% 
    filter(projectCountry %in% c("Mexico", "Colombia", "Peru"))

entityData <- entityClean %>% 
    filter(entityCountry %in% c("Mexico", "Colombia", "Peru"))

countrySummary <-  tibble(country = c("Mexico", "Colombia", "Peru")) %>% 
    rowwise() %>% 
    mutate(nProjects = n_distinct(projectData[projectData$projectCountry == country, ]$projectName), 
           nEntities = n_distinct(entityData[entityData$entityCountry == country, ]$entityName), 
           avgProjectPayment = mean(projectData[projectData$projectCountry == country, ]$projectPayment)/1000000,
           avgEntityPayment = mean(entityData[entityData$entityCountry == country, ]$entityPayment)/1000000,
           totProjectPayment = sum(projectData[projectData$projectCountry == country, ]$projectPayment)/1000000,
           totEntityPayment = sum(entityData[entityData$entityCountry == country, ]$entityPayment)/1000000)

names(countrySummary) <- c("Country", "No. of Projects", "No. of Entities", "Avg. Proj Payment (in mill. USD", "Avg. Entity Payment (in mill. USD)", "Total Proj Payment (in mill. USD)", "Total Entity Payment (in mill. USD)")

kable(countrySummary, digits = 2) %>% 
    kable_styling(bootstrap_options = c("striped", "hover")) 
    
```
The interesting thing to note here is that while Colombia has lesser number of projects than Mexico and Peru it has a higher total payment amount, hinting that the per project payments are higher for Colombia.

Next lets plot the distribution of overall payments across project and entity payments.

```{r, echo = FALSE, message=FALSE, fig.align='center'}
joinData <- entityData %>% 
    select(country = entityCountry, payment = entityPayment) %>% 
    mutate(type = "Entity")
combinedData <- projectData %>% 
    select(country = projectCountry, payment = projectPayment) %>% 
    mutate(type = "Project") %>% 
    bind_rows(., joinData)

combinedData %>% 
    ggplot(., aes(payment, x = factor(country))) +
    geom_boxplot() +
    facet_wrap(~type, ncol = 1) +
    theme_tufte() +
    labs(y = "Payments (in USD)", title = "Payments to projects and entities") +
    coord_flip() +
    theme(
                axis.text.x = element_text(size = 8),
                axis.title.x = element_text(size = 10),
                axis.title.y = element_blank(),
                legend.position = "none",
                plot.title = element_text(size = 18),
                strip.text.x = element_text(size = 10)
        )
```

The dots on the chart above indicate outliers (greater than the 90th percentile of payment values for each payment type). Both projects and entity payments seem to have large right tails indicating that a few large payments dominate the total revenues that are generated. 

##By Payment Types
We would need to look at the specific splits by payment types to see if these are spread equally across different categories. First lets look at the summary of the number of payments, the average amounts and the total by each payment type for each country. From now, on we are only considering the project payments since the data is similar for both.

```{r, echo = FALSE, message=FALSE, results='asis'}
table1 <- projectData %>% 
    select(Country = projectCountry, paymentType, projectPayment) %>% 
    group_by(Country, paymentType) %>% 
    summarise(`Total Payment` = sum(as.numeric(projectPayment)/1000000, na.rm=T), 
              `Avg. Payment` = mean(as.numeric(projectPayment)/1000000, na.rm=T),
              `No. of Payments` = n()) %>% 
    ungroup() %>% 
    gather(measure, value, 3:5) %>% 
    arrange(paymentType, Country, measure) %>% 
    spread(paymentType, value) %>% 
    group_by(Country, measure) %>% 
    mutate(`Avg./Total`= ifelse(measure == "Avg. Payment", round(mean(c(Bonuses, Fees, `Payments for infrastructure improvements`, `Production entitlements`, Royalties, Taxes), na.rm = T), 2), round(sum(c(Bonuses, Fees, `Payments for infrastructure improvements`, `Production entitlements`, Royalties, Taxes), na.rm = T), 2))) %>% 
    ungroup() %>% 
    select(-Country)


kable(table1, caption = "Payments by Type (in million USD)", digits = 2) %>% 
    kable_styling(bootstrap_options = c("striped", "hover")) %>% 
    column_spec(1, width = "60em", italic = T) %>% 
    column_spec(2:7, width = "30em") %>% 
    column_spec(8, width = "50em") %>% 
    group_rows("Colombia", 1, 2, 3, label_row_css = "background-color: #666; color: #fff;") %>% 
    group_rows("Mexico", 4, 5, 6, label_row_css = "background-color: #666; color: #fff;") %>% 
    group_rows("Peru", 7, 8, 9, label_row_css = "background-color: #666; color: #fff;")
```

As expected the average payment values for Royalties and Taxes for Colombia are higher than the other two countries. Now lets plot the total amount of payments by type for each country. 

```{r, echo=FALSE, fig.align="center"}
projectData %>%
    ggplot(., aes(x = projectCountry, y = projectPayment, fill = paymentType)) +
    geom_bar(stat = "identity", position = "dodge") +
    theme_tufte() +
    scale_fill_viridis(discrete = T) +
    labs(y = "Total Payments (in USD)") +
    theme(
                axis.text.x = element_text(size = 8),
                axis.title.x = element_blank(),
                axis.title.y = element_text(size = 10),
                legend.position = "top",
                legend.title = element_blank(),
                plot.title = element_text(size = 18),
                strip.text.x = element_text(size = 10)
        )
```
The three countries have similar profiles, however, there are some interesting differences. While all of them seem to rely significantly on taxes, Colombia and Peru rely on Royalties while Mexico seems to rely more heavily on Fees. Lets take a closer look at the spread of the payment values for each country and payment type.

```{r, echo=FALSE, fig.align="center"}
paymentLimit <- 5000000
projectData %>% 
    ggplot(., aes(projectPayment, x= factor(paymentType))) +
    geom_point(aes(colour = projectPayment)) +
    geom_hline(yintercept = paymentLimit, colour = "blue", alpha = 0.6) +
    scale_color_viridis(alpha = 0.8, option = "magma", direction = -1) +
    facet_wrap(~projectCountry, ncol = 1) +
    theme_tufte() +
    labs(y = "Project Payments (in USD)", title = "Project Payments by Type") +
    coord_flip() +
    theme(
                axis.text.x = element_text(size = 8),
                axis.title.x = element_text(size = 10),
                axis.title.y = element_blank(),
                legend.position = "none",
                plot.title = element_text(size = 18),
                strip.text.x = element_text(size = 10)
        )
```

The blue line is set at 5 million USD for reference. As can be seen, most of the large payments except for a few Royalty payments in Colombia are in the taxes category. This would indicate that if these large payments were removed we'd see a more even distribution of payments across different types.

##Large Payments
There are 24 payment entries out of the 293 entries for these countries that are greater than 5 million USD. These 24 payements constite approximately 72% of the total revenue generated in these three countries. In other words, 8% of the payments generate 72% of the revenues.

Lets plot the total payment values by type after removing payment values that are greater than 5 million USD.

```{r, echo=FALSE, fig.align="center"}
sum(projectData[projectData$projectPayment > paymentLimit, ]$projectPayment)/sum(projectData$projectPayment)
projectData %>%
    filter(projectPayment < paymentLimit) %>% 
    ggplot(., aes(x = projectCountry, y = projectPayment, fill = paymentType)) +
    geom_bar(stat = "identity", position = "dodge") +
    theme_tufte() +
    scale_fill_viridis(discrete = T) +
    labs(y = "Total Payments (in USD)") +
    theme(
                axis.text.x = element_text(size = 8),
                axis.title.x = element_blank(),
                axis.title.y = element_text(size = 10),
                legend.position = "top",
                legend.title = element_blank(),
                plot.title = element_text(size = 18),
                strip.text.x = element_text(size = 10)
        )
```

Removing project payment values greater than 5 million USD seems to have had a big impact on taxes. It seems that most of the bigger payment values for all these countries were in the taxes category. Lets see if this bears out if we see the spread of payments.

Lets look at these large payments in detail. The table below shows the payments that were more than 10 million USD

```{r, echo=FALSE, results='asis'}
projectData %>% 
    filter(projectPayment > paymentLimit * 2) %>% 
    mutate(projectPayment = projectPayment/1000000) %>% 
    select(Company = reportingCompany, Country = projectCountry, Project = projectName, Type = paymentType, `Value (in USD)` = projectPayment) %>% 
    arrange(Country, desc(`Value (in USD)`)) %>% 
    kable(., digits = 2) %>% 
    kable_styling(bootstrap_options = c("striped", "hover"))
```

Lets look at the splits for these large projects

```{r echo=FALSE, fig.align="center"}
projectData %>%
    filter(projectPayment > paymentLimit) %>% 
    ggplot(., aes(x = projectCountry, y = projectPayment, fill = paymentType)) +
    geom_bar(stat = "identity", position = "dodge") +
    theme_tufte() +
    scale_fill_viridis(discrete = T) +
    labs(y = "Total Payments (in USD)") +
    theme(
                axis.text.x = element_text(size = 8),
                axis.title.x = element_blank(),
                axis.title.y = element_text(size = 10),
                legend.position = "top",
                legend.title = element_blank(),
                plot.title = element_text(size = 18),
                strip.text.x = element_text(size = 10)
        )
```
As expected, majority of the larger payments were in taxes across all three countries.

##Things to Explore

1. Try different payment value buckets (small, medium, large) and see if there are any patterns to observe.

