---
title: "Data stuff for the workshop"
output: 
  html_notebook:
    number_sections: yes
    toc: true
    toc_depth: 3
    toc_float: true
    theme: cosmo
    highlight: kate
    smart: true
    code_folding: hide
date: '`r Sys.Date()`'
---

##Data issue questions

1. What is the RP converage by commodity
2. How many companies pay taxes to countries that are not the project country or reporting country?
3. Proportion of payments by joint ventures
4. Compare ownership on SNL against whether they are filing on RP
5. Single asset companies list

##Summary
Lets load the latest data and functions.

```{r, error=FALSE, message=FALSE, results='hide'}
rm(list = setdiff(ls(), c("sourceData", "projectData", "entityData", "currentLookup", "currentSecondary", "snlData")))
library(tidyverse); library(ggthemes); library(kableExtra); library(knitr); library(viridis); library(readxl); library(stringr); library(googlesheets)

if(!exists("sourceData")){ ##so that we only run on first load
    sourceData <- read_excel("../../Data Cleaning/Payment Data Pipeline/data/Current website data/allDataNew.xlsx", sheet = "sources")
    projectData <- read_excel("../../Data Cleaning/Payment Data Pipeline/data/Current website data/allDataNew.xlsx", sheet = "projects")
    entityData <- read_excel("../../Data Cleaning/Payment Data Pipeline/data/Current website data/allDataNew.xlsx", sheet = "entities")
}

```

Load and process SNL data.
```{r, error=FALSE, message=FALSE, results='hide'}
if(!exists("snlData")){
    snlCmdtyLocData <- read_excel("../../Data Cleaning/Secondary Data/Projects/Data/Secondary data sources/snl_all_comdty_location.xls", sheet = 1) %>% 
    mutate_at(16:17, as.character) %>% 
    bind_rows(., read_excel("../../Data Cleaning/Secondary Data/Projects/Data/Secondary data sources/snl_all_comdty_location_inactive.xls", sheet = 1)) %>% 
    select(-(`Property Name`))

    snlOwnership <- read_excel("../../Data Cleaning/Secondary Data/Projects/Data/Secondary data sources/snl_ownership.xls", sheet = 1) %>% 
        bind_rows(., read_excel("../../Data Cleaning/Secondary Data/Projects/Data/Secondary data sources/snl_ownership_inactive.xls", sheet = 1)) %>% 
        select(-(`Property Name`))
    
    snlFinancing <- read_excel("../../Data Cleaning/Secondary Data/Projects/Data/Secondary data sources/snl_financing.xls", sheet = 1) %>% 
        bind_rows(., read_excel("../../Data Cleaning/Secondary Data/Projects/Data/Secondary data sources/snl_financing_inactive.xls", sheet = 1)) %>% 
        gather(key = "round", value = "amt", -(1:2)) %>% 
        mutate(amt = as.numeric(amt)) %>% 
        group_by(`Property ID`) %>% 
        summarise(totalFinance = sum(amt, na.rm = T)) %>% 
        mutate(totalFinance = ifelse(totalFinance == 0, NA, totalFinance))
    
    snlProduction <- read_excel("../../Data Cleaning/Secondary Data/Projects/Data/Secondary data sources/snl_production.xls", sheet = 1) %>% 
        bind_rows(., read_excel("../../Data Cleaning/Secondary Data/Projects/Data/Secondary data sources/snl_production_inactive.xls", sheet = 1)) %>% 
        gather(key = "name", value = "value", -(1:2)) %>%
        mutate(value = as.numeric(value)) %>% 
        filter(!is.na(value)) %>%
        arrange(`Property ID`, desc(value)) %>% 
        group_by(`Property ID`) %>% 
        filter(row_number() < 3) %>% 
        summarise(avgProduction = mean(value, na.rm = T))
    
    snlData <- left_join(snlCmdtyLocData, snlOwnership, by = "Property ID") %>% 
        left_join(., snlFinancing, by = "Property ID") %>% 
        left_join(., snlProduction, by = "Property ID")
    
    rm(snlCmdtyLocData, snlOwnership, snlFinancing, snlProduction)
}

```

Load the lookup tables and secondary data
```{r, error=FALSE, message=FALSE, results='hide'}
if(!exists("currentLookup")){
    gs_auth("/Users/hsubhash/Documents/GitHub/resource-project-data/token.rds")
    currentLookup <- gs_title("Current Lookup") %>% 
        gs_read(ws = "Current Matched Lookup", verbose = F)
    currentSecondary <- gs_title("Current Secondary") %>% 
        gs_read(ws = "Secondary Data", verbose = F)

    currentSecondary <- currentSecondary %>% 
        mutate(projectType = case_when(
        str_detect(str_to_lower(projectType), "corporate") ~ "Corporate",
        str_detect(str_to_lower(projectType), "oil") ~ "Oil & Gas",
        str_detect(str_to_lower(projectType), "mining") ~ "Mining",
        str_detect(str_to_lower(projectType), "other") ~ "Other"
        )) %>% 
        mutate(locationType = case_when(
        str_detect(str_to_lower(locationType), "exact") ~ "Exact",
        str_detect(str_to_lower(locationType), "capital") ~ "Regional Capital",
        str_detect(str_to_lower(locationType), "nearest town") ~ "Nearest Town",
        str_detect(str_to_lower(locationType), "estimated") ~ "Estimated"
        )) %>% 
        mutate(commodity = ifelse(commodity == ",,,", NA, commodity)) %>% 
        filter(paste0(projectCountry, finalName) %in% paste0(currentLookup$projectCountry, currentLookup$finalName)) %>% 
        select(-sources, -notes, -duplicates)
}

```


## Joint Ventures
These are projects that are reported by multiple companies
```{r}
jvProjects <- projectData %>% 
    mutate(projectPayment = as.numeric(projectPayment)) %>% 
    filter(!is.na(projectPayment) & projectPayment != 0) %>% 
    left_join(., currentLookup %>% 
                  select(1:4), by = c("projectCountry", "reportingCompany", "projectName" = "rawName")) %>% 
    mutate(projectName = ifelse(is.na(finalName), projectName, finalName)) %>% 
    group_by(projectCountry, reportingCompany, projectName, paymentType, reportYear) %>% 
    summarise(projectPayment = sum(projectPayment, na.rm = T)) %>% 
    group_by(projectCountry, projectName, reportYear) %>% 
    filter(n_distinct(reportingCompany) > 1) %>% 
    filter(projectName != "Corporate") %>% 
    ungroup() %>% 
    left_join(., currentSecondary, by = c("projectCountry", "projectName" = "finalName")) %>% 
    filter(!is.na(projectCountry) | projectCountry != "x") %>% 
    filter(!str_detect(str_to_lower(projectName), "alberta|columbia|other|general|gulf of mexico|attribu|operated|entity|exploration|united states of")) %>% 
    filter(projectName != "x" & projectName != "Bc") %>% 
    arrange(projectCountry, projectName, reportingCompany, paymentType)
    


```

```{r}
jvProjects <- jvProjects %>% 
    group_by(projectCountry, reportingCompany, projectName) %>% 
    mutate(totalPay = sum((projectPayment), na.rm = T)) %>% 
    ungroup() %>% 
    mutate(proportionPay = projectPayment/totalPay) %>% 
    write_csv(., "Temp Data/JV Data/jv.csv")
```








