---
title: "Exploratory analysis for RP data workshop"
output: 
  html_notebook:
    code_folding: hide
    highlight: kate
    number_sections: yes
    theme: cosmo
    toc: yes
    toc_float: yes
date: '`r Sys.Date()`'
---




```{r, error=FALSE, message=FALSE, echo=FALSE, results='hide'}
rm(list = ls())
library(tidyverse); library(ggthemes); library(kableExtra); library(knitr); library(viridis); library(readxl); library(stringr)

# projectData <- read_csv("../../Data Cleaning/Secondary Data/Projects/Data/Temp Files/projectsFinal.csv")
# #snlData <- read_csv("../../Data/Secondary Data/Projects/Data/Temp Files/secondaryProjectData.csv")
currentProjectData <- read_csv("../../Data Cleaning/Payment Data Pipeline/data/Current website data/allProject.csv") %>% 
    select(projectCountry, reportingCompany, projectName, paymentType, projectPayment)

snlCmdtyLocData <- read_excel("../../Data Cleaning/Secondary Data/Projects/Data/Secondary data sources/snl_all_comdty_location.xls", sheet = 1) %>% 
    mutate_at(16:17, as.character) %>% 
    bind_rows(., read_excel("../../Data Cleaning/Secondary Data/Projects/Data/Secondary data sources/snl_all_comdty_location_inactive.xls", sheet = 1)) %>% 
    select(-(`Property Name`))

snlOwnership <- read_excel("../../Data Cleaning/Secondary Data/Projects/Data/Secondary data sources/snl_ownership.xls", sheet = 1) %>% 
    bind_rows(., read_excel("../../Data Cleaning/Secondary Data/Projects/Data/Secondary data sources/snl_ownership_inactive.xls", sheet = 1)) %>% 
    select(-(`Property Name`))

snlFinancing <- read_excel("../../Data Cleaning/Secondary Data/Projects/Data/Secondary data sources/snl_financing.xls", sheet = 1) %>% 
    bind_rows(., read_excel("../../Data Cleaning/Secondary Data/Projects/Data/Secondary data sources/snl_financing_inactive.xls", sheet = 1)) %>% 
    gather(key = "round", value = "amt", -(1:2)) %>% 
    mutate(amt = as.numeric(amt)) %>% 
    group_by(`Property ID`) %>% 
    summarise(totalFinance = sum(amt, na.rm = T)) %>% 
    mutate(totalFinance = ifelse(totalFinance == 0, NA, totalFinance))

snlProduction <- read_excel("../../Data Cleaning/Secondary Data/Projects/Data/Secondary data sources/snl_production.xls", sheet = 1) %>% 
    bind_rows(., read_excel("../../Data Cleaning/Secondary Data/Projects/Data/Secondary data sources/snl_production_inactive.xls", sheet = 1)) %>% 
    gather(key = "name", value = "value", -(1:2)) %>%
    mutate(value = as.numeric(value)) %>% 
    filter(!is.na(value)) %>%
    arrange(`Property ID`, desc(value)) %>% 
    group_by(`Property ID`) %>% 
    filter(row_number() < 3) %>% 
    summarise(avgProduction = mean(value, na.rm = T))

snlData <- left_join(snlCmdtyLocData, snlOwnership, by = "Property ID") %>% 
    left_join(., snlFinancing, by = "Property ID") %>% 
    left_join(., snlProduction, by = "Property ID") %>% 
    select(snlPropID = "Property ID",
           startUpYear = "Actual Start Up Year",
           primaryCommodity = "Primary Commodity",
           commodityList = "List of Commodities",
           mineType1 = "Mine Type 1",
           mineType2 = "Mine Type 2",
           mineType3 = "Mine Type 3",
           lat = "Latitude\n\n(degrees)",
           lon = "Longitude\n\n(degrees)",
           stage = "Development Stage",
           status = "Activity Status",
           reservesResources = "Reserves & Resources: Ore Tonnage\n\n(tonnes)",
           avgProduction,
           operator = "Operator Name",
           operatorValue = "Operator Total Enterprise Value\n\n(Reported)",
           owner = "Owner Name",
           ownerValue = "Owner Total Enterprise Value\n\n(Reported)"
           ) %>% 
    mutate(reservesResources = as.numeric(reservesResources))


secondaryData <- read_csv("../../Data Cleaning/Secondary Data/Projects/Data/Temp Files/Secondary files/Current Secondary.csv")

currentLookup <- read_csv("../../Data Cleaning/Secondary Data/Projects/Data/Temp Files/Lookup files/Current Lookup.csv")

rm(snlCmdtyLocData, snlOwnership, snlFinancing, snlProduction)
```

##Look at SNL data

```{r}

```






```{r}
snlTemp <- snlData 

allData <- projectData %>% 
    left_join(., snlTemp, by = "snlPropID") %>% 
    filter(!is.na(snlPropID))
```



```{r, error=FALSE, message=FALSE, echo=FALSE, results='hide'}
snlTemp <- snlData %>% 
    select(projectCountry = "Country Name",
           snlPropID = "Property ID",
           startUpYear = "Actual Start Up Year",
           primaryCommodity = "Primary Commodity",
           stage = "Development Stage",
           status = "Activity Status",
           reserves = "Reserves & Resources: Ore Tonnage\n\n(tonnes)",
           avgProduction) %>% 
    mutate(reserves = as.numeric(reserves))
```

##Comparison of SNL data with RP data
This file uses the entire list of SNL projects that are available on the portal that consists of **`r nrow(snlData)` projects** and those projects that are in the RP database that could be mapped to SNL, which consists of **`r n_distinct(projectData$snlPropID)` projects**.

###Development stage and activity status
Lets first look at the number of projects by different development stages. There is a predominance of projects in the `Operating` category in RP data when compared to the SNL data, which not surprising given that we would expect most of the payments to happen for projects that are in the `Operating` stage of their life cycles. There also seems to be a disproportionate representation of projects in the bottom half of the table that are in stages such as `Prefeas/Scoping` or `Expansion` compared to their overall shares in the SNL data.

```{r, message=FALSE}
snlSum <- snlTemp %>% 
    group_by(stage) %>% 
    summarise(`No. of SNL Projects` = n()) %>% 
    arrange(desc(`No. of SNL Projects`)) %>% 
    bind_rows(., tibble(stage = "Total", `No. of SNL Projects` = nrow(snlData)))

projectSum <- projectData %>% 
    select(snlPropID) %>% 
    unique() %>% 
    left_join(., snlTemp, by = "snlPropID") %>%
    group_by(stage) %>% 
    summarise(`No. of RP Projects` = n()) %>% 
    arrange(desc(`No. of RP Projects`)) %>% 
    bind_rows(., tibble(stage = "Total", `No. of RP Projects` = n_distinct(projectData$snlPropID)))

left_join(snlSum, projectSum, by = "stage") %>% 
    kable(., digits = 2) %>% 
    kable_styling(bootstrap_options = c("striped", "hover"))
```


Next lets look at the activity types. Again, the numbers seem to be in line with what we would expect. Most of the projects in the RP data are in the `Active` category. Interestingly enough there are a significant number of projects that are `Inactive` that received payments over the last two years. Lets take a closer look at these.
```{r, message=FALSE}
snlSum <- snlTemp %>% 
    group_by(status) %>% 
    summarise(`No. of SNL Projects` = n()) %>% 
    arrange(desc(`No. of SNL Projects`)) %>% 
    bind_rows(., tibble(status = "Total", `No. of SNL Projects` = nrow(snlData)))

projectSum <- projectData %>% 
    select(snlPropID) %>% 
    unique() %>% 
    left_join(., snlTemp, by = "snlPropID") %>%
    group_by(status) %>% 
    summarise(`No. of RP Projects` = n()) %>% 
    arrange(desc(`No. of RP Projects`)) %>% 
    bind_rows(., tibble(status = "Total", `No. of RP Projects` = n_distinct(projectData$snlPropID)))

left_join(snlSum, projectSum, by = "status")%>% 
    kable(., digits = 2) %>% 
    kable_styling(bootstrap_options = c("striped", "hover"))
```
The table below shows the `Inactive` projects that recieved payments. Most of these seem to have been in the `Exploration` stage and likely were closed after unsuccessful exploration.
```{r, message=FALSE}
projectData %>% 
    select(snlPropID) %>% 
    unique() %>% 
    left_join(., snlTemp, by = "snlPropID") %>% 
    filter(status == "Inactive") %>% 
    group_by(stage) %>% 
    summarise(`No. of RP Projects` = n()) %>% 
    arrange(desc(`No. of RP Projects`)) %>% 
    kable(., digits = 2) %>% 
    kable_styling(bootstrap_options = c("striped", "hover"))
```

Next lets look at the commodity data
###Compare commodities
The table below shows the different commodites and their respective payouts. Overall, `Gold` seems to be the predominate commodity in both SNL and RP data. However, the total payment for `Iron` projects is approximately 3 times that for all the gold projects. This is despite there being about 20 times more number of gold projects. As is evident from the production values, the average production from gold mine is about 400 times lower than that from a gold mine.

**Note:** Table only show commodities that have more than 10 projects present in the RP data. Additionally, production values were not reported for U308 and diamond projects in SNL.
```{r, message=FALSE}
snlSum <- snlTemp %>% 
    group_by(commodity) %>% 
    summarise(`No. of SNL Projects` = n()) %>% 
    arrange(desc(`No. of SNL Projects`)) 

projectSum <- projectData %>% 
    group_by(snlPropID) %>% 
    summarise(totPay = sum(projectPayment, na.rm = T)) %>%
    ungroup() %>% 
    left_join(., snlTemp, by = "snlPropID") %>%
    group_by(commodity) %>% 
    summarise(`No. of RP Projects` = n(), 
              `Tot. Payment recieved` = sum(totPay, na.rm = T), 
              `Avg payment (per mine)` = mean(totPay, na.rm = T),
              `Std Dev of Payments` = sd(totPay, na.rm = T), 
              `Avg. Reserves (in tonnes)` = mean(ifelse(reserves < 0, 0, reserves), na.rm = T),
              `Avg. Production (in tonnes)` = mean(ifelse(avgProduction == 0, NA, avgProduction), na.rm = T)) %>% 
    arrange(desc(`No. of RP Projects`))

left_join(snlSum, projectSum, by = "commodity") %>% 
    filter(!is.na(`No. of RP Projects`) & `No. of RP Projects` > 10) %>% 
    kable(., digits = 0) %>% 
    kable_styling(bootstrap_options = c("striped", "hover"))
```


```{r, message=FALSE}
normalize <- function(vec){
    maxVal <- max(vec, na.rm = T)
    
    out <- numeric()
    
    for(i in 1:length(vec)){
        out[i] <- (vec[i]/maxVal)*100
    }
    out
}

left_join(snlSum, projectSum, by = "commodity") %>% 
    filter(!is.na(`No. of RP Projects`) & `No. of RP Projects` > 10) %>% 
    select(1, 3, 4, 5, 8) %>% 
    mutate_at(2:5, normalize) %>% 
    gather(key = "variable", value = "value", -1) %>% 
    filter(!is.nan(value)) %>% 
    mutate(variable = factor(variable, levels = c("No. of RP Projects", "Tot. Payment recieved", "Avg payment (per mine)", "Avg. Production (in tonnes)"))) %>% 
    ggplot(., aes(x = commodity, y = value, fill = variable)) +
    geom_bar(stat = "identity", position = "dodge") +
    scale_fill_viridis(discrete = T, direction = -1)
```


```{r}
chartData <- projectData %>% 
    left_join(., snlTemp, by = "snlPropID") %>% 
    mutate(startUpYear = as.numeric(ifelse(startUpYear == "NA", NA, startUpYear))) %>% 
    mutate(age = (2018 - (startUpYear)))%>%
    filter(!is.na(age)) %>% 
    mutate(ageCat = factor(cut(age, breaks = c(0, 2, 5, 10, 20, 30, 1000)), labels = c("0 to 5", "5 to 10", "10 to 15", "15 to 20", "20 to 30", "Greater than 30"))) %>% 
    group_by(commodity, age) %>% 
    filter(projectPayment < quantile(projectPayment, 0.95) & projectPayment > 0 & age < 40) %>% 
    mutate(avPay = mean(projectPayment), countAge = n()) %>% 
    ungroup() %>% 
    group_by(commodity) %>% 
    mutate(countCom = n()) %>% 
    filter(countCom > 30) %>%
    ungroup() %>% 
    mutate(commodity = paste0(commodity, " (",countCom, " obs.) "))


ggplot() + 
    geom_line(data = chartData %>% 
                  distinct(commodity, age, avPay), aes(x = age, y = avPay)) +
    geom_point(data = chartData %>% 
                  distinct(commodity, age, avPay), aes(x = age, y = avPay), alpha = 0.4, colour = "blue") +
    geom_text(data = chartData %>% 
                  distinct(commodity, age, avPay, countAge), aes(x = age, y = avPay, label = countAge), vjust = -0.25, size = 2.5) +
    geom_smooth(data = chartData, aes(x = age, y = projectPayment), method = 'loess', level = 0.9) +
    facet_wrap(~commodity, ncol = 1, scales = "free_y") +
    labs(x = "Age of the mine (in years)", y = "Avg. payment received (in USD)") +
    theme_tufte() +
    theme(legend.title = element_blank(), 
          strip.text.x = element_text(size = 14),
          legend.position = "top",
          title = element_text(size = 16),
          legend.text = element_text(size = 12)
          )


# chartData %>% 
#     group_by(commodity) %>% 
#     filter(n() > 15) %>% 
#     ungroup() %>% 
#     filter(projectPayment < quantile(projectPayment, 0.9)) %>% 
#     ggplot(., aes(x = commodity, y = projectPayment)) + 
#     geom_jitter(aes(color = factor(ageCat))) +
#     geom_boxplot(alpha = 0.3, colour = "orange", outlier.colour =  NA) +
#     coord_flip() +
#     scale_color_viridis(discrete = T, alpha = 0.5) + 
#     theme_tufte() +
#     theme(legend.title = element_blank(), 
#           legend.position = "top",
#           axis.title.x = element_blank(),
#           title = element_text(size = 16),
#           legend.text = element_text(size = 12)
#           )
```

































```{r}
snlData %>% 
    select(projectCountry = "Country Name", 
           snlPropID = "Property ID", 
           commodity = "Primary Commodity",
           stage = "Development Stage",
           status = "Activity Status") %>% 
    filter(!is.na(commodity)) %>% 
    summarise(nProjects = n_distinct(snlPropID), nActive = )

    
    kable(., digits = 2) %>% 
    kable_styling(bootstrap_options = c("striped", "hover"))    
```



T


###Reserves by country

```{r}
chartData <- snlData %>% 
    select(projectCountry = "Country Name", snlPropID = "Property ID", commodity = "Primary Commodity", reserves = "Reserves & Resources: Ore Tonnage\n\n(tonnes)", reservesVal = "In-Situ Value: Reserves & Resources\n\n(Reported)") %>% 
    mutate_at(4:5, as.numeric)
```


```{r}
chartData %>% 
    filter(!is.na(reservesVal) & reservesVal > 0 & reservesVal < quantile(reservesVal, 0.9, na.rm = T)) %>% 
    group_by(commodity) %>% 
    filter(n() > 100) %>% 
    ungroup() %>% 
    ggplot(., aes(reservesVal)) +
    geom_density() +
    theme_tufte() +
    facet_wrap(~commodity, scales = "free_y", ncol = 1)
```

##Comparison between snl data and RP data

```{r}
snlSum <- snlData %>% 
    select(projectCountry = "Country Name", 
           snlPropID = "Property ID", 
           commodity = "Primary Commodity",
           stage = "Development Stage",
           status = "Activity Status") %>% 
    filter(!is.na(commodity))
   
```


###Commodity vs payment

```{r}

    
chartData <- projectData %>% 
    group_by(snlPropID) %>% 
    summarise(totPay = sum(projectPayment, na.rm = T)) %>%
    ungroup() %>% 
    left_join(., snlData %>% 
                  select(snlPropID = "Property ID", 
                         commodity = "Primary Commodity",
                         reserves = "Reserves & Resources: Ore Tonnage\n\n(tonnes)"),
              by = "snlPropID") %>%
    mutate(reserves = as.numeric(reserves)) %>% 
    group_by(commodity) %>% 
    summarise(nProjects = n_distinct(snlPropID), allPay = sum(totPay, na.rm = T), avgPay = mean(totPay, na.rm = T), sdPay = sd(totPay, na.rm = T),totReserves = sum(ifelse(reserves > 0, reserves, 0), na.rm = T))
```

MAKE THIS CHART



###Reserves vs payment

```{r}
chartData <- projectData %>% 
    mutate(snlPropID = as.character(snlPropID)) %>% 
    group_by(snlPropID) %>% 
    summarise(totPay = sum(projectPayment, na.rm = T)) %>%
    left_join(., snlData %>% 
                  select(snlPropID, reserves = "Reserves & Resources: Ore Tonnage\r\r(tonnes)", 
                         reservesVal = "In-Situ Value: Reserves & Resources\r\r(Reported)",
                         commodity = "Primary Commodity"), 
              by = "snlPropID") %>% 
    mutate(reserves = as.numeric(reserves), reservesVal = as.numeric(reservesVal)) %>% 
    mutate(reserves = ifelse(reserves == -2147483648, NA, reserves))

chartData %>% 
    filter(!is.na(reservesVal) & reservesVal < quantile(reservesVal, 0.9, na.rm = T) & totPay < quantile(totPay, 0.9) & totPay > 0) %>% 
    ggplot(., aes(x = reservesVal, y = totPay)) + 
    geom_point(aes(color = factor(commodity))) + 
    geom_smooth(method = "lm") +
    theme_tufte()

```

