---
title: "Summary of UK mandatory disclosures"
author:
    name: Hari Subhash
    affiliation: Natural Resource Governance Institute
output: 
  html_notebook:
    number_sections: yes
    toc: true
    toc_depth: 3
    toc_float: true
    theme: cosmo
    highlight: kate
    smart: true
    code_folding: hide
date: "`r format(Sys.time(), '%d %B %Y')`"
---

```{r, echo=FALSE, message=FALSE}
rm(list = ls())
library(tidyverse); library(lubridate)
ukSources <- read_csv("../../resource-project-data/Data Cleaning/Payment Data Pipeline/data/Current website data/allSource.csv") %>% 
    filter(reportingCountry == "United Kingdom of Great Britain and Northern Ireland")
ukProjects <- read_csv("../../resource-project-data/Data Cleaning/Payment Data Pipeline/data/Current website data/allProject.csv") %>% 
    filter(sourceID %in% ukSources$sourceID)
ukEntities <- read_csv("../../resource-project-data/Data Cleaning/Payment Data Pipeline/data/Current website data/allEntity.csv") %>% 
    filter(sourceID %in% ukSources$sourceID)
```

There are **`r nrow(ukSources)`** mandatory disclosure reports that have been disclosed in UK from **`r n_distinct(ukSources$reportingCompany)`** companies from **`r min(year(ymd(ukSources$reportStartDate)))`** to **`r max(year(ymd(ukSources$reportStartDate)))`**. The companies made a total payment of **`r sum(ukProjects$projectPayment, na.rm = T)/1e9`** billion USD to **`r n_distinct(ukProjects$projectCountry)`** different countries.

The chart below shows the break down of the number of reports over the different years.

```{r}

ggplot(ukSources %>%
        group_by(year(ymd(ukSources$reportStartDate))) %>% 
        summarise(`Number of Reports` = n_distinct(sourceID)) %>% 
        select(year = `year(ymd(ukSources$reportStartDate))`, 2),
       aes(x = as.factor(year), y = `Number of Reports`)) +
    geom_bar(stat = "identity", width = 0.5)
```

