---
title: "Introduction to Resource projects"
date: '`r Sys.Date()`'
author: "Hari Subhash"
output:
  revealjs::revealjs_presentation:
    includes:
       in_header: header.html
    css: assets/styles.css
    fig_width: 7
    fig_height: 6
    fig_caption: true
    theme: "sky"
    highlight: "haddock"
    self_contained: false
    reveal_plugins: ["notes", "search", "zoom"]
---
```{r, message=FALSE, echo=FALSE}
library(tidyverse); library(emo)

##load current website data
allSources <- read_csv("../../resource-project-data/Data Cleaning/Payment Data Pipeline/data/Current website data/allSource.csv")
allProjects <- read_csv("../../resource-project-data/Data Cleaning/Payment Data Pipeline/data/Current website data/allProject.csv")
allEntities <- read_csv("../../resource-project-data/Data Cleaning/Payment Data Pipeline/data/Current website data/allEntity.csv")

# Create the external file
img <- htmltools::img(src = knitr::image_uri(file.path("/Users/hsubhash/Documents/GitHub/RP Outreach/assets/images/RP_logo.png")), 
               alt = 'logo', 
               style = 'position:absolute; top:20px; right:1%; padding:10px;z-index:200; width:100px')

htmlhead <- paste0('<link href="https://fonts.googleapis.com/css?family=Open+Sans|Roboto" rel="stylesheet">', '
<script>
    document.write(\'<div class="logos">',img,'</div>\')
</script>
')

readr::write_lines(htmlhead, path = "header.html")

```

##{data-background-image="https://cdn.rawgit.com/NRGI/RP-Outreach/f2da482a/assets/images/home_page.png"}

## Objective

- Give an overview of [resourceprojects.org](https://resourceprojects.org/) and mandatory disclosure data.



## Agenda

> - What is [resourceprojects.org](https://resourceprojects.org/)?
> - Summary of the data
> - Accessing the data on resourceprojects.org


# Summary
This R notebook generates all the files and analysis used for the Introduction to RP presentation.


First lets gather some basic summary information.

Number of reporting jurisdictions
```{r}
# print(paste("The total payment reported is ", round(sum(allProjects$projectPayment, na.rm = T)/1e9, 1), "billion USD"))
# print(paste("The number of reporting reports are ", n_distinct(allSources$sourceID)))
# print(paste("The number of reporting jurisdictions are ", n_distinct(allSources$reportingCountry)))
# print(paste("The number of reporting companies are ", n_distinct(allSources$reportingCompany)))
# print(paste("For the years", min(allSources$reportYear), "to", max(allSources$reportYear)))

```
Number of reporting companies.


##Test


##Test 2


##Normal

